Fail Faster

### 2025-03-23 paper/paper.tex
chart

\begin{figure}[h]
  \includegraphics[scale=0.28]{bind.png}
  \caption{Relationship between the number of binds---a proxy for reliance on generator library abstractions---and speedup from staging, showing that the two relate linearly.}
  \label{fig:binds}
\end{figure}

Results are summarized in \figref{fig:binds}. The left plot shows a clear linear 
relationship between performance benefit from staging and number of calls 
to \texttt{bind}. The right plot shows a more bimodal distribution, but still validates our intuition that
generators that sample more also stand to gain more from optimized \rand libraries.
These results suggest that staging and \rand optimizations address distinct and complimentary roles in generator performance: the former aids combinator-heavy generators, while the latter benefits sampling-heavy ones. To support diverse workloads, generator libraries should combine both techniques.

@@

As shown in \figref{fig:scala}, \scalaname achieves an even greater performance
gain---purely from staging---over ScalaCheck than \camlname does over \bq. In particular,
the single-pass BST strategy is $4.89-6.51\times$ faster, the boolean list generator is $6.86-13.41\times$ faster,
and STLC is $5.43-9.70\times$ faster. This difference arises from
ScalaCheck's representation of generators as functions of type \texttt{size ->
seed -> Option[A]}: each generator combinator must construct and then
pattern-match on these \texttt{Option} values, introducing significant boxing
and unboxing overhead at each step.  Staging eliminates this
overhead.

>>>>

\begin{figure}[h]
  \includegraphics[scale=0.28]{bind.png}
  \caption{Left: Speedup from staging (compared to baseline \bq) versus the number of \texttt{bind} calls.
  Right: Speedup from \csplitmix (compared to staging alone) versus the number of random samples.}
  \label{fig:binds}
\end{figure}

\figref{fig:binds} summarizes our results. The left plot shows a clear linear 
relationship between performance benefit from staging and number of calls 
to \texttt{bind}. The right plot shows a more bimodal distribution, but still validates our intuition that
generators that sample more also stand to gain more from optimized \rand libraries.
These results suggest that staging and \rand optimizations address distinct and complimentary roles in generator performance: the former aids combinator-heavy generators, while the latter benefits sampling-heavy ones. To support diverse workloads, generator libraries should combine both techniques.

@@

As shown in \figref{fig:scala}, \scalaname achieves an even greater performance
gain---purely from staging---over ScalaCheck than \camlname does over \bq. In particular,
the single-pass BST strategy is $4.89-6.51\times$ faster, the boolean list generator is $6.86-13.41\times$ faster,
and STLC is $5.43-9.70\times$ faster. This difference arises from
ScalaCheck's representation of generators as functions of type \texttt{size ->
seed -> Option[A]}: each generator combinator must construct and then
pattern-match on these \texttt{Option} values, introducing significant boxing
and unboxing overhead at each step.  Staging eliminates this
overhead. These results show that staging generalizes well across languages.
